---
title: "R Notebook"
output: html_notebook
---
Smallest multiple
```{r}

```
  
Problem 5
2520 is the smallest number that can be divided by each of the numbers from 1 to 10 without any remainder.

What is the smallest positive number that is evenly divisible by all of the numbers from 1 to 20?

I thought this was going to be a small number, which was why I tried this method... (got upto 20 million before I stopped it)

Slight cheat by copying this from hackerrank.com
M is divisible by all numbers from 1 to N iff M is equal to the product of all prime powers p^k where p is prime and divides M, and p < N, and k is the largest integer such that p^k < N. For example, 2520 = 2^3 * 3^2 * 5 * 7.

Where N = 1:20
ans = M


```{r}

l = 1

a = F

while (a == F){

  boolean = rep(F, 20)
  
  for (i in c(1:20)){
  
    if (l %% i == 0) {
      boolean[i] = T
    }  
    
  }
  
  if (all(boolean == T)) {
    ans = l
    a = T
    return(ans)
  }
  
  if (l %% 5000 == 0){print(cat(l, " ", boolean, " sum is...", sum(boolean)))}
  l = l+1
}

```

Least Common multiple
Nice video here.... https://www.youtube.com/watch?v=z5vlj06Ex0U

Requires my prime number function...
```{r}

prime_numbers <- function(num) {
  
  if (num < 2) return(NULL)
  
  primes_boolean <- rep(T, num)
  
  primes_boolean[1] <- F
  
  for(i in seq(num)) {
   
     if (primes_boolean[i]) {
      
       i_sq <- i * i
       
      if (i_sq > num) return(which(primes_boolean))
      
       primes_boolean[seq(i_sq, num, by=i)] <- FALSE
    }
  }
}
```


```{r}

arr = seq(1,20)

LCM = function(arr) {
  
  primes = prime_numbers(max(arr))
  
  
}
```

