---
title: "R Notebook"
output: html_notebook
---
Even Fibonacci numbers

Problem 2
Each new term in the Fibonacci sequence is generated by adding the previous two terms. By starting with 1 and 2, the first 10 terms will be:

1, 2, 3, 5, 8, 13, 21, 34, 55, 89, ...

By considering the terms in the Fibonacci sequence whose values do not exceed four million, find the sum of the even-valued terms.

Fibonacci function, using the closed form inside a loop
```{r}

# Function to find Nth fibonacci number, and form an array through recursion
fib = function(x) {
  
  # Golden ratio
  g = (1 + 5**.5) / 2
  
  ls = integer()
  
  for (i in 1:x){
    
    j = (g**i - (1-g)**i) / 5**.5
    
    ls[i] = j
    
    # Once we've gone over the input number, cut the last one off and break the loop
    if (j > x){
      ls = ls [ls<x]
      break
    }
  } 
  
  ls
}

```

Subset out the even numbers
```{r}
# subset the even numbers
# (modulo needs to take an integer input, not numeric)
a = ls[as.integer(ls) %% 2 == 0]

```

Sum the new list
```{r}
ans = sum(a)
```


```{r}
ls = fib(4000000)
```

