---
title: "R Notebook"
output: html_notebook
---
Lattice paths
  
Problem 15
Starting in the top left corner of a 2×2 grid, and only being able to move to the right and down, there are exactly 6 routes to the bottom right corner.

How many such routes are there through a 20×20 grid?

--- 

it always takes four moves (for 2x2), for an i x i grid, it'll take 2i moves

so for 20 x 20 it'll be 40 moves

half of them must be down and half right

every series of moves has an 'opposite' which is the same but reversed

can be represented by a 1:40 chain of 1s and 0s, i think, each one being a move
  total of the chain must equal 20
  
  so how many combinations of 0 and 1, 40 in length, make up 20?
  
  if it's 4 in length:
  0011
  0101
  0110
  1001
  1010
  1100 = 6
  
  6 in length
  
  000111
  001011
  001101
  001110
  010011
  010101
  010110
  011001
  011010
  011100
  100011
  100101
  100110
  101001
  101010
  101100
  110001
  110010
  110100
  111000 = 20

Total possible combinations must be 2^i

Maybe the total, over i, is the number of combinations that equal i/2?

for i = 4
combos c = 2^4 = 16
combos that equal i/2 = c/i = 16/4 = 4 

so that's not right...

for i = 4, there's 4 combos that = 1
  5 wher c = 2
  6 wher c = 3
  1 wher c = 4
  
for i = 6, there's 6 where c = 1
  15 where c = 2
  

```{r}
2^4
```

I did a lot of trial runs on paper working through the example with 1s and 0s

then did the same with a box and arrows

ended up researching the CENTRAL BINOMIAL COEFFICIENT

which is:

(2n/n) = 2n!/((n!)^2)

so for n = 4 (which is the case for a 4 x 4 box)

= 8! / 4! ^2 = ( 8 x 7 x 6 x 5 x 4! ) / ( 4! x 4!) = (cancel the 4!s) = 8x7x6x5 / 4x3x2x1

so the question becomes, what is 40!/20!^2

Using a factorial function

```{r}

cbc = function(n){
  x = factorial(2*n)/(factorial(n))^2
  return(x)
}
```



```{r}
ans = cbc(20)

ans
```