---
title: "Euler 3"
output: html_notebook
---
Largest prime factor
 
Problem 3
The prime factors of 13195 are 5, 7, 13 and 29.

What is the largest prime factor of the number 600851475143 ?



System:

Step 1 create vector of primes up to 1,000,000
Step 2 - go through the list of primes, dividing num by each until mod = 0
Step 3 - Store that number in a vector and update num to the output of the division
Step 4 - When new_num = 0, return the largest number in the new vector

function to calculate prime numbers upto a given number
```{r}
prime_numbers = function(top) {

 candidates = c(2:top)
 
 prime_nums = c()
 
 for (i in c(2:top)) {
  
   if (any(candidates == i)) {
    
    prime_nums = c(prime_nums, i)
    
    candidates = c(candidates[(candidates %% i) != 0], i)
    
   }
 }
 
 return(prime_nums)
 
 }

```



```{r}
num = 13195
new_num = num
primes = prime_numbers(num)
t = length(primes)
factors = integer()
p = 1

for (i in 1:t){
  if (new_num %% primes[i] == 0) {
    print(cat("adding ", primes[i], " to factors"))
    factors[p] = primes[i]
    p = p+1
    new_num = new_num/primes[i]
  }
  if (new_num == 1) {
    ans = max(factors)
    print(cat("answer is ", ans))
    break
  }
}

```






now use the function to find prime factors
```{r}

factors = function(n) {
  
  max_primes = 500

  primes = prime_numbers(max_primes)
  
  fctrs = c()
  
  j = 1
  
  while (n != 0){
  
    for (i in c(j:max_primes)) {
      
      print(cat("i = ", i,", j is", j, ", fctrs is", fctrs))
      
      if (n %% primes[i] == 0){
    
        fctrs = c(fctrs, primes[i])
      
        n = n/primes[i]
        j = j+1
    
        break
      }
      
     
    }
    
    
  }
  
  return(fctrs)
}
```


