---
title: "Euler 3"
output: html_notebook
---
Largest prime factor
 
Problem 3
The prime factors of 13195 are 5, 7, 13 and 29.

What is the largest prime factor of the number 600851475143 ?



System:

Step 1 - create vector of primes up to 1,000,000
Step 2 - go through the list of primes, dividing num by each until mod = 0
Step 3 - Store that number in a vector and update num to the output of the division
Step 4 - When new_num = 0, return the largest number in the new vector

function to calculate prime numbers upto a given number


from interweb... sieve of erastothenes
```{r}


sieve <- function(n) {
  if (n < 2) return(NULL)
  a <- rep(T, n)
  a[1] <- F
  for(i in seq(n)) {
    if (a[i]) {
      j <- i * i
      if (j > n) return(which(a))
      a[seq(j, n, by=i)] <- F
    }
  }
}

```



My version of the Sieve of Erastothenes...

Step 1 - if input is less than 2, break the function
Step 2 - make a boolean vector of all possible numbers upto [num], default all as TRUE
Step 3 - make '1' FALSE
Step 4 - loop through all numbers upto [num]
  if the number at i is not already marked as false
    square the number at i
    if the square is less than [num]
      mark all multiples of the square of i, upto [num] as false
    else (square of i is greater than [num]), return the index of all the TRUE values in the boolean vector
      

```{r}

prime_numbers <- function(num) {
  
  if (num < 2) return(NULL)
  
  primes_boolean <- rep(T, num)
  
  primes_boolean[1] <- F
  
  for(i in seq(num)) {
   
     if (primes_boolean[i]) {
      
       i_sq <- i * i
       
      if (i_sq > num) return(which(primes_boolean))
      
       primes_boolean[seq(i_sq, num, by=i)] <- FALSE
    }
  }
}

```


My new code, based on a more simple stepwise iteration than previsou fancy attempt


```{r}

num = 600851475143
new_num = num
primes = prime_numbers(sqrt(num))
t = length(primes)
factors = integer()
p = 1

for (i in 1:t){
  if (new_num %% primes[i] == 0) {
    print(cat("adding ", primes[i], " to factors"))
    factors[p] = primes[i]
    p = p+1
    new_num = new_num/primes[i]
  }
  if (new_num == 1) {
    ans = max(factors)
    print(cat("answer is ", ans))
    break
  }
}

```




