---
title: "R Notebook"
output: html_notebook
---
Highly divisible triangular number
  
Problem 12
The sequence of triangle numbers is generated by adding the natural numbers. So the 7th triangle number would be 1 + 2 + 3 + 4 + 5 + 6 + 7 = 28. The first ten terms would be:

1, 3, 6, 10, 15, 21, 28, 36, 45, 55, ...

Let us list the factors of the first seven triangle numbers:

 1: 1
 3: 1,3
 6: 1,2,3,6
10: 1,2,5,10
15: 1,3,5,15
21: 1,3,7,21
28: 1,2,4,7,14,28
We can see that 28 is the first triangle number to have over five divisors.

What is the value of the first triangle number to have over five hundred divisors?

---

Ok so what's the relationship between nth trianglular number and number of factors?

Could loop, incrementally add one to the list of triangular numbers, and check the number of divisors...

number of factors can be found be using prime factorisation (take code from earlier notebook?), and then using :

N = Xa × Yb × Zc

Num of prime factros = (a+1)(b+1)(c+1)

where N is the number to be factorised, X, Y and Z are the prime numbers and a, b and c are their respective powers.


Steps
-----

1. Find next triangular number (t)
2. Calculate primes upto sqrt of t (use a while?)
3. Prime factorisation of t, using vecotr of primes
Find  Num of factors with = (a+1)(b+1)(c+1)
if num of factors = >500, return t


```{r}



# prime factors function
pr_fac = function(n) {
  x = c() # array of prime factors
  r = n # remainder
  i = 2 # index
  
  while (r > 1) { # while remainder is greater than 0
    
    if (r %% i == 0) { # if remainder is divisble by i
      
      x = c(x,i) # add i to the prime factors vector
      r = r/i # update remainder
      i = 1 # reset i to 1 (will get 1 more added below)
      
    }
    
    i = i + 1
    
  }
  return(x)
}

# function to find the numebr of factors (inc non-primes)
factors = function(n){

  prime_factors = pr_fac(n) # uses function pr_fac to return vector of prime factors (may include duplicates)
 
  tbl = table(prime_factors) # make a frequency table of prime factors (elements are the count of each prime, so are equivalent to the exponents)

  exponents = as.vector(tbl) # extract the exponents
 
  n_factors = prod(exponents + 1) # Num of prime factors = (a+1)(b+1)(c+1)

  return(n_factors)
}

# loop through triangle numbers until f > 500
t = 1
i = 1
f = 0

while (f < 500){
  i = i + 1
  t = t + i
  f = factors(t)
}

t

```




-----
Prime numbers upto N function from euler 3 

```{r}



# prime_numbers
# returns a vector of all primes less than n

prime_numbers <- function(n) {
  
  if (n < 2) return(NULL)
  
  primes_boolean <- rep(T, n)
  
  primes_boolean[1] <- F
  
  for(i in seq(n)) {
   
     if (primes_boolean[i]) {
      
       i_sq <- i * i
       
      if (i_sq > n) return(which(primes_boolean))
      
       primes_boolean[seq(i_sq, n, by=i)] <- FALSE
    }
  }
}


# sum_nth
# returns the sum of primes upto nth prime

sum_nth = function(n){
  primes = c(2,3) # array of primes
  q = 5           # next candidate to be checked for primacy
  while(q<n){
      if (all(q %% primes[1:sqrt(q)] != 0)){ 
        primes = c(primes, q)
        # cat("\n", "q is: ", q, "\n", "primes looks like: ", head(primes, 5), "... ", tail(primes, 5), "\n", "sum is: ", sum)
      }
      
      # add two to check the next odd number
      q = q + 2
  }
  return(sum(primes))
}


# prime_factors
# returns a vector of prime factors of n, and a vector of their exponents

# steps

# divide by first prime
# if remainder is 0, add first prime to factors and update remainder
# if remainder is not zero, calculate next prime
# q = last in primes + 2
# if q is not divisble by any in primes, add it to primes
# else, add 2 to q


# 
# prime_factors = function(n){
#   
#   primes = 2          # array of primes
#   q = 3               # next candidate to be checked for primacy
#   factrs = 0          # array of prime factors
#   expos = 0           # vector of prime exponents
#   remainder = n
#   
#   while(remainder > 0) {
#     
#     while (any(remainder %% primes == 0)){                        # if remainder can be divided by any number in prime...
#       
#       for (i in length(primes)){                                  # go through the list of primes
#         
#         if (remainder %% primes[i] == 0) {                      # if remainder can be divided by primes[i] 
#         
#           factrs = c(factrs, primes[i])                           # add ith primes to list of prime factors
#           # expos[length(expos)] = expos[length(expos)] + 1               # add 1 to the last exponent in the list
#           
#           remainder = remainder / q                             # update remainder
#         
#         }
#       }
#       
#     }
#     
#     
#     q = q + 2
#     
#       # calculate the next prime 
#         
#     
#     while (any(q %% primes == 0)){     # while q is not prime
#       q = q+2                                     # add two to check the next odd number
#     }
#       
#     primes = c(primes, q)
#       
#       
#           # cat("\n", "q is: ", q, "\n", "primes looks like: ", head(primes, 5), "... ", tail(primes, 5), "\n", "sum is: ", sum)
#     
#       
#                                               
#     }
#     return(factrs, expos)
# }  

# try again!

pr_fac = function(n) {
  x = c() # array of prime factors
  r = n # remainder
  i = 2 # index
  
  while (r > 1) { # while remainder is greater than 0
    
    if (r %% i == 0) { # if remainder is divisble by i
      
      x = c(x,i) # add i to the prime factors vector
      r = r/i # update remainder
      i = 1 # reset i to 1 (will get 1 more added below)
      
    }
    
    i = i + 1
    
  }
  return(x)
}

# Ok this one works!!!
  
```
  
  
Code from Euler 3 to find the largest prime factor (relies on function above)


```{r}

num = 600851475143
new_num = num
primes = prime_numbers(sqrt(num))
t = length(primes)
factors = integer()
p = 1

for (i in 1:t){
  if (new_num %% primes[i] == 0) {
    print(cat("adding ", primes[i], " to factors"))
    factors[p] = primes[i]
    p = p+1
    new_num = new_num/primes[i]
  }
  if (new_num == 1) {
    ans = max(factors)
    print(cat("answer is ", ans))
    break
  }
}

```

Could utilise my improved primes function... (but this is for the sum of primes up to n)

```{r}


=
```



---