---
title: "R Notebook"
output: html_notebook
---
10001st prime
  
Problem 7
By listing the first six prime numbers: 2, 3, 5, 7, 11, and 13, we can see that the 6th prime is 13.

What is the 10 001st prime number?

---

Surely just use the prime generator function from euler-3...

actually realised we need to return the 10001th, but the function gives us primes upto 10001...#

Can modify by

```{r}
prime_numbers <- function(num) {
  
  if (num < 2) return(NULL)
  
  primes_boolean <- rep(T, num)
  
  primes_boolean[1] <- F
  
  for(i in seq(num)) {
   
     if (primes_boolean[i]) {
      
       i_sq <- i * i
       
      if (i_sq > num) return(which(primes_boolean))
      
       primes_boolean[seq(i_sq, num, by=i)] <- FALSE
    }
  }
}
```

Using Wilson's theorem from wikipedia

nth_prime = ((n_factorial mod (n+1))/n)*(n-1)+2

First needs a factorial function...

```{r}
factorial = function(f){
  ans = 1
  for (i in 1:f){
    ans = ans * i
  }
  return(ans)
}

```
Tried that with 1000 and answer was 'inf'... so suspect 10000 won't work!

A web factorial calculator says the largest it will go is 170!

---

From stackexchange...

we have the following asymptotic bound on the nth prime pn:

n ln n + n(ln ln n − 1) < pn < n ln n + n ln ln n for n ≥ 6

So we know roughly where it is, but that's not very helpful...

```{r}
n = 10000
lower2 = n * log(n) * n + n * (log(log(n)) - 1)
upper2 = n * log(n) * n + n * log(log(n))

diff(c(upper1, lower2))

```



Another approach is to store all primes upto N by iterating through all the integers <N

Only need to divide each one by the primes already in the list

And only odd numbers

Starting at 3

(but it's not 'upto N' because that doesn't give us the Nth prime)

So changed now to keep iterating until lenth of primes = N

Buut... still takes a loooong time

So maybe should use the limits above to shorten time

```{r}

primes2 = function(n){
  ## define bounds for the nth prime
  lower = n * log(n) * n + n * (log(log(n)) - 1)
  upper = n * log(n) * n + n * log(log(n))
  
  ##check that the bounds don't overlap with previous or next prime
  prev = n-1
  prev_upper = prev * log(prev) * prev + prev * log(log(prev))
  nxt = n+1
  nxt_lower = nxt * log(nxt) * nxt + nxt * (log(log(nxt)) - 1)
  if (lower<prev_upper){return("overlap with previous window")}
  if (upper>nxt_lower){return("overlap with next window")}
  
  ## use the sqrt of the upper bound as upper limit for calculating primes to be used as factors
  sqrt_upper = sqrt(upper)
  
  ## make list of primes upto sqrt_upper
  primes = c(2,3)
  i = 5
  while(length(primes)<sqrt_upper){
    if (all(i %% primes != 0)){
      primes = c(primes,i)
      print(cat("adding ", i, " to primes"))
    }
    i = i+2 
    print(i)
  }
  
  ## then use modified version of my old primes() format to calculate primes between lower and upper (should only be one) 
  
  lower = floor(lower)
  upper = ceiling(upper)
  arr = c(lower:upper)
  
  arr_boolean <- rep(T, diff(c(lower, upper)))
  
  for(i in arr) {
   
    if (arr_boolean[i]) {
      j = lower + i
       
      if (j > upper){ return(arr[which(arr_boolean)])}
      
      else if (any(arr[i] %% primes == 0)) {arr_boolean[i] <- FALSE}
    }
  }

  
}


N = 10001
primes2(N)


```

Ok this system is stupid... I'm currently calculating ALL primes upto sqrt(upper), when i shoudl be generating the next (smallest) prime
and marking all numbers in the array that are divisible as false

steps
step1 - generate list from lower to upper [list = 10:20]
step4 - start a list of primes p, starting with 2
step3 - generate boolean index from lower to upper [boo = T, T, T, T, T, T, T, T, T, T, T, T, T, T, T, T, T, T, T, T]
step4 - define q as the last element in primes + 1
step4 - mark all boolean's with an index greater than [lower] and where it's index plus [lower] is divisible by q, as false
step - find next prime and add it to p
    while tail(primes) is less than sqrt(upper),
      q = tail(primes)+1
      if all(q %% primes ==0){ primes = c(primes,q) }
      break the while
      
      
```{r}
nth_prime = function(n){

  ## start my list of primes
  primes = c(2)
  
  ## Wilson's theorem to set boundaries for nth prime (and round them with floor/ceiling)
  lower = floor(n * (log(n) + log(log(n)) - 1))
  upper = ceiling(n * (log(n) + log(log(n))))
  
  arr = c(lower:upper)
  
  boo = rep(T, length(arr))
  
  while (tail(primes,1) < upper){
  
    ## make elements of boo corresponding to elements of arr that are divisible by q as false
    boo[which(arr %% tail(primes, 1) == 0)] = FALSE
  
    ## next potential prime
    q = tail(primes, 1) + 1
  
    ## check if it's prime
    while(T){
      if (all(q %% primes != 0)){ 
        primes = c(primes, q)
        break
        }
      else q = q + 1
    }
  }
return(arr[boo])
}
```
      
      