---
title: "R Notebook"
output: html_notebook
---
Largest palindrome product
    
Problem 4

A palindromic number reads the same both ways. The largest palindrome made from the product of two 2-digit numbers is 9009 = 91 × 99.

Find the largest palindrome made from the product of two 3-digit numbers.


Thoughts...

There can't be very many palindromic numbers which are the product of two 3-digit numbers.
```{r}
100*100
a = 100 * 100
```

```{r}
999*999
b = 999*999
```

```{r}
b - a
```
... possible answers

So what if I start at 999 * 999 and work backwards till I find a palindrome...
```{r}

for (i in seq(999^2, 100^2)) {

  if (i %% 500 == 0){ print(i)}
  
  a = strsplit(as.character(i), "")
  b = lapply(strsplit(as.character(i), ""), rev)
  
  if (all(unlist(a) == unlist(b))){
    ans = i
    
    print(cat(i, " is palindromic"))
    if (any(q == ans)){
      return(ans)
      break
    }
    else {print("but not a product of two 3-digit numbers")}
  }
}

```



So that works, but the answer it returns (997799) might not be a product of two 3-digit numbers...
... how to check that...


I could make a list of all product of all numbers and check it off...

(this bit needs to be run before the code above)

```{r}
q = data.frame()

for (i in c(1:900)) {
  for (j in c(1:900)) {
    q[i,j] = (i+99)*(j+99)
   
  }
 
}
```



