---
title: "R Notebook"
output: html_notebook
---
10001st prime
  
Problem 7
By listing the first six prime numbers: 2, 3, 5, 7, 11, and 13, we can see that the 6th prime is 13.

What is the 10 001st prime number?

---

Surely just use the prime generator function from euler-3...

actually realised we need to return the 10001th, but the function gives us primes upto 10001...#

Can modify by

```{r}
prime_numbers <- function(num) {
  
  if (num < 2) return(NULL)
  
  primes_boolean <- rep(T, num)
  
  primes_boolean[1] <- F
  
  for(i in seq(num)) {
   
     if (primes_boolean[i]) {
      
       i_sq <- i * i
       
      if (i_sq > num) return(which(primes_boolean))
      
       primes_boolean[seq(i_sq, num, by=i)] <- FALSE
    }
  }
}
```

Using Wilson's theorem from wikipedia

nth_prime = ((n_factorial mod (n+1))/n)*(n-1)+2

First needs a factorial function...

```{r}
factorial = function(f){
  ans = 1
  for (i in 1:f){
    ans = ans * i
  }
  return(ans)
}

```
Tried that with 1000 and answer was 'inf'... so suspect 10000 won't work!

A web factorial calculator says the largest it will go is 170!

---

From stackexchange...

we have the following asymptotic bound on the nth prime pn:

n ln n + n(ln ln n − 1) < pn < n ln n + n ln ln n for n≥6


```{r}
n = 10000
lower2 = n * log(n) * n + n * (log(log(n)) - 1)
upper2 = n * log(n) * n + n * log(log(n))

diff(c(upper1, lower2))

```

